---
title: "Copy of Old WHO MMR Group Project Notebook"
output: pdf_document
html_document:
html_notebook: default
pdf_document: 
keep_md: yes
---
#this notebook can be edited w/o worrying about losing work

```{r}
library(tidyr)
library(tidyverse)
library(readxl)
library(dplyr)
library(devtools)
library(usethis)
library(ggplot2)

country_info <- read_excel("country list_ 26 March 2019.xlsx")
mmr_est_unrounded <- read.csv("mmr_unrounded.csv")
live_birth_projections <- read_excel("wpp2019_Births-TFR-GFR-Female1549.xlsx")
regional_groupings <- read.csv("regional_groupings_20190910_la_ac.csv")

load_all("WHOmmr3")
```

#Data Cleaning
```{r}
country_info <- country_info %>%
  select(`Code`, `ISO_Numeric_Code_CODE`, `Title`) 
  
mmr_est_unrounded <- mmr_est_unrounded %>%
  filter(`bound` == "point", `year` >= 2010) %>%
  select(-c(`q`, `perc`, `bound`)) %>%
  mutate("MMR" = `value`*100000)

mmr_est_unrounded_pwider <- mmr_est_unrounded %>%
  select(-c(`value`)) %>%
  pivot_wider(names_from = `year`, values_from = `MMR`)
  
regional_groupings <- regional_groupings %>%
  select(`ISOCode`, `Country.x`, `sdg_1`)

live_birth_projections2 <- live_birth_projections %>% 
  filter(Year == 2030)
colnames(live_birth_projections2)[colnames(live_birth_projections2)=="Location"] <- "name"
```

#BAU ARR Function
Calculates the business as usual ARR for all countries (adds column to provided tibble)
```{r}
bau_arr(mmr_est_unrounded_pwider, 2010, 2017)
```

#MMR Projection Function (One Country)
Calculates the projected MMRs from 2016 to 2030 for a single, specified country
```{r}
mmr_country_projections(mmr_est_unrounded_pwider, 2010, 2017, iso_code = "AFG")
```

#MMR Projection Function (All Countries)
Calculates the projected MMRs from 2016 to 2030 for all countries
```{r}
mmr_allcountries_projections(mmr_est_unrounded_pwider, 2010, 2017)
```

#SDG MMR Calculation, Categorization, and Adjustment
Calculates the projected MMR for a specified year based on a global_arr, then:
- categorizes the country's MMR as high or acceptable 
- readjusts the high MMR countries' MMR to meet the SDG requirements
```{r}
global_arr<- mean(bau_arr(mmr_est_unrounded_pwider, 2010, 2017)$`arr`)

SDG_MMR_2030(mmr_est_unrounded_pwider, 2030, global_arr)
```

#Regional MMR Summary Table
Calculates the projected MMRs from 2016 to 2030 based on region
```{r}
#some data cleaning
colnames(country_info)[colnames(country_info)== "Code"] <- "ISOCode"

countries_and_regions <- country_info %>% 
  left_join(regional_groupings, by = c("ISOCode" = "ISOCode")) %>%
  select(-c(`Country.x`))

colnames(countries_and_regions)[colnames(countries_and_regions)== "ISO_Numeric_Code_CODE"] <- "ISONum"


#function
mmr_regional_global_summarize(mmr_est_unrounded_pwider, countries_and_regions, live_birth_projections2)
```

#Plot Regional Summaries
```{r}
mmr_regional_global_summaries <-
mmr_regional_global_summarize(mmr_est_unrounded_pwider, countries_and_regions, live_birth_projections2)

mmr_graph <- mmr_regional_global_summaries %>%
  rename(`2016`="MMR in 2016") %>%
  rename(`2017`="MMR in 2017") %>%
  rename(`2018`="MMR in 2018") %>%
  rename(`2019`="MMR in 2019") %>%
  rename(`2020`="MMR in 2020") %>%
  rename(`2021`="MMR in 2021") %>%
  rename(`2022`="MMR in 2022") %>%
  rename(`2023`="MMR in 2023") %>%
  rename(`2024`="MMR in 2024") %>%
  rename(`2025`="MMR in 2025") %>%
  rename(`2026`="MMR in 2026") %>%
  rename(`2027`="MMR in 2027") %>%
  rename(`2028`="MMR in 2028") %>%
  rename(`2029`="MMR in 2029") %>%
  rename(`2030`="MMR in 2030") %>%
pivot_longer(2:16,names_to="period",values_to="mmr")
mmr_graph %>%
ggplot() + geom_point(aes(x=period,y=mmr, color=Region)) +
  geom_line(mapping = aes(x=period, y=mmr, group=Region, color = Region))
```

#Make a graphical summary of regional summaries?