---
title: "Recoding For Numerical Output Only"
output: html_notebook
---

#Read-in Data
```{r}
library(tidyr)
library(tidyverse)
library(readxl)
#library(dplyr) #maybe don't need
library(devtools)
library(usethis)

country_info <- read_excel("country list_ 26 March 2019.xlsx")
mmr_est_unrounded <- read.csv("mmr_unrounded.csv")
live_birth_projections <- read_excel("wpp2019_Births-TFR-GFR-Female1549.xlsx")
regional_groupings <- read.csv("regional_groupings_20190910_la_ac.csv")

```

#Data Cleaning
```{r}
country_info <- country_info %>%
  select(`Code`, `ISO_Numeric_Code_CODE`, `Title`) 
  
mmr_est_unrounded <- mmr_est_unrounded %>%
  filter(`bound` == "point", `year` >= 2010) %>%
  select(-c(`q`, `perc`, `bound`)) %>%
  mutate("MMR" = `value` * 100000)

mmr_est_unrounded_pwider <- mmr_est_unrounded %>%
  select(-c(`value`)) %>%
  pivot_wider(names_from = `year`, values_from = `MMR`)
  
regional_groupings <- regional_groupings %>%
  select(`ISOCode`, `Country.x`, `sdg_1`)

live_birth_projections2030 <- live_birth_projections %>% 
  filter(Year == 2030)
colnames(live_birth_projections2030)[colnames(live_birth_projections2030) == "Location"] <- "name"
```


#Calc BAU ARR
- original parameters: mmr tibble, time1, time2
- UN-HARDCODE specified years when they're the name of the columns
- returns: vector of observed ARR for specified time period for all countries
```{r}
#function
cba <- function(mmr_pivotwider_tibble, start_year, end_year) {
  length_years = end_year - start_year
  arr = (-1/(length_years))*log(mmr_pivotwider_tibble[["2017"]]/mmr_pivotwider_tibble$`2010`)
  #return(cbind(mmr_pivotwider_tibble, arr))
  return(arr)
}

#example
cba(mmr_est_unrounded_pwider, 2010, 2017)

#is this a vector w/ entries?
```


#MMR Projections (for one country)
- original parameters: mmr tibble, time1, time2, country_iso_code
- UNHARDCODE the number 2015 and the column name
- returns: a single column tibble w/ all the projected MMRs
```{r}
#function
mcp <- function(mmr_pivotwider_tibble, mmr_start_year = 2016, mmr_end_year = 2030, bau_start_year = 2010, bau_end_year = 2017, iso_code){

  years <- seq(mmr_start_year, mmr_end_year)
  num_years <- length(years)
  bau_tibble <- calc_bau_arr(mmr_pivotwider_tibble, bau_start_year, bau_end_year) %>% 
    filter(iso == iso_code)
  projection <- rep(NA,num_years) #What does this do?
  projection <- bau_tibble$`2015` * exp(-(bau_tibble$`arr`) * (years-2015))
  return(projection)
  }

#example
mcp(mmr_est_unrounded_pwider, 2016, 2030, 2010, 2017, iso_code = "AFG")
```


SDG MMR Calculation, Categorization, and Adjustment
#This function currently calculates the MMR of 2030 based on a global arr from a simple calculation
- original parameters: mmr tibble, single year, global arr
- UNHARDCODE the number 2015 and the column name; the 2010 and 2017 in calc_bau_arr() function call
- SOLVE the issues w/ multiple tibbles and the resulting dependencies
- returns: tibble w/ countries'observed arr, mmr projections during time period, 2030 projection, each country's MMR category, each country's adjusted SDG goal
- Do we want this to simply add a column to the tibble passed thru asa an argument? Or just the calculated SDG MMR? Or maybe that plus the country codes/names? 
```{r}
#function
SM2030 <- function(mmr_pivotwider_tibble, arr, years_to_project) {
  sdg_beforefix140 <- mmr_pivotwider_tibble[[1]] * exp(-global_arr * (years_to_project))
 sdg_afterfix140 <- ifelse(sdg_beforefix140 >140, 140, sdg_beforefix140)
  return(sdg_afterfix140)
}

global_arr <- mean(calc_bau_arr(mmr_est_unrounded_pwider, 2010, 2017)$`arr`)

#example
mmr2015 <- select(mmr_est_unrounded_pwider, `2015`)
SM2030(mmr2015, global_arr, 15)

```

#Squared Diff
- for optimization purposes in conjunction w/ function above
```{r}
#function
  sd <- function(global_arr, mmr2015_country,
                         births2030_country, nproject){
  
  testmmr <- SDG_MMR_2030(mmr2015_country, global_arr, nproject)
  # testbirth <- births2030_country %>%
  # right_join(mmr_est_unrounded_pwider, by = c("name" = "name")) %>%
  #   select(`Births`)
  
  mb1 <- filter(testmmr*testbirth, `SDG_projection`>100)
  mb2 <- filter(testmmr*testbirth, `SDG_projection`<=100)
  b1 <- filter(testbirth, `Births`>500)
  b2 <- filter(testbirth, `Births`<=500)
  mmr_global <- (sum(mb1) + sum(mb2)) / (sum(b1) + sum(b2)) 
  (mmr_global - 70)^2

}

#example - FIXME
births2030 <- live_birth_projections2030 %>%
  right_join(mmr_est_unrounded_pwider, by = c("name" = "name")) %>%
    select(`Births`)
sd(global_arr, mmr2015, births2030, 15)
  
```