---
title: "WHO MMR Working Notebook"
output: html_notebook
---

#Read-in Data
```{r}
library(tidyr)
library(tidyverse)
library(readxl)
library(devtools)
library(usethis)

Rfiles <- list.files(file.path(paste0(getwd(),"/R/")), ".R")
Rfiles <- Rfiles[grepl(".R", Rfiles)]
sapply(paste0(paste0(getwd(),"/R/"), Rfiles), source)

country_info <- read_excel("country list_ 26 March 2019.xlsx")
mmr_est_unrounded <- read.csv("mmr_unrounded.csv")
live_birth_projections <- read_excel("wpp2019_Births-TFR-GFR-Female1549.xlsx")
regional_groupings <- read.csv("regional_groupings_20190910_la_ac.csv")

```

#Data Cleaning
```{r}
country_info <- country_info %>%
  select(`Code`, `ISO_Numeric_Code_CODE`, `Title`) %>%
  rename(ISOCode = `Code`, ISONum = `ISO_Numeric_Code_CODE`)
  
mmr_est_unrounded <- mmr_est_unrounded %>%
  filter(`bound` == "point", `year` >= 2010) %>%
  select(-c(`q`, `perc`, `bound`)) %>%
  mutate("MMR" = `value` * 100000)

mmr_est_unrounded_pwider <- mmr_est_unrounded %>%
  select(-c(`value`)) %>%
  pivot_wider(names_from = `year`, values_from = `MMR`)
  
regional_groupings <- regional_groupings %>%
  select(`ISOCode`, `Country.x`, `sdg_1`)

live_birth_projections2030 <- live_birth_projections %>% 
  filter(Year == 2030) %>%
  select(-c(`Year`)) %>%
  rename(name = 'Location') 
```

#Calc BAU ARR
- original parameters: mmr tibble
- UN-HARDCODE specified years when they're the name of the columns
- returns: tibble w/ added observed ARR for specified time period

```{r}
calc_bau_arr(mmr_est_unrounded_pwider, 2010, 2017)

#new function only has bau arr and iso code columns
cba(mmr_est_unrounded_pwider)


#Question:
# the function considers the length of the time period to be 7 years but it's actually 8 years b/c we're counting 2010 in the period
```


#MMR Projections (for one country)
- original parameters: mmr tibble, time1, time2, country_iso_code
- UNHARDCODE the number 2015 and the column name
- returns: a single column tibble w/ all the projected MMRs

```{r}
mmr_country_projections(mmr_est_unrounded_pwider, 2016, 2030, 2010, 2017, iso_code = "AFG")
    
#Recoded
mcp(mmr_est_unrounded_pwider, "AFG")

```

#MMR Projections (all countries)
- original parameters: mmr tibble, time1, time2
- UNHARDCODE column names
- RENAME "col" and "namecountry2" to something more intuitive
- returns: a single column tibble w/ all the projected MMRs
- is there a way to remove the column numbers of the final tibble?

```{r}
mmr_allcountries_projections(mmr_est_unrounded_pwider, 2010, 2017)

#new function does not have arr column
macp(mmr_est_unrounded_pwider)

```

#SDG MMR Calculation, Categorization, and Adjustment
#This function currently calculates the MMR of 2030 based on a global arr from a simple calculation
- original parameters: mmr tibble, single year, global arr
- UNHARDCODE the number 2015 and the column name; the 2010 and 2017 in calc_bau_arr() function call
- SOLVE the issues w/ multiple tibbles and the resulting dependencies
- returns: tibble w/ countries'observed arr, mmr projections during time period, 2030 projection, each country's MMR category, each country's adjusted SDG goal
- Do we want this to simply add a column to the tibble passed thru asa an argument? Or just the calculated SDG MMR? Or maybe that plus the country codes/names? 

```{r}
global_arr <- mean(cba(mmr_est_unrounded_pwider)$`arr`)

mmr2015 <- mmr_est_unrounded_pwider %>% 
  rename(MMR2015 = `2015`) %>%
  select(`MMR2015`)

get_mmr2030_sdg_projections(mmr2015, global_arr)

##

```

#Squared Diff
- for optimization purposes in conjunction w/ function above
```{r}
squared_diff(global_arr, mmr2015, live_birth_projections2030, 15)
```

#Get ARR SDG Target 
- original parameters: mmr2015_country (vector w/ MMR for 2015 for all countries), births2030_country (vector w/ births for 2030 for all countries), nproject (number of years to project (2030 - 2015))
- UNHARDCODE the specified years each argument holds (eg make it possible for users to use different years for projection bases and different year ranges, etc)
- consider making the values that the initial arguments hold different and actually calculated in the function body
- returns: tibble w/ global(?) arr sdg target for the cumulative period of 15 years
#Pseudo code
#This should feed into the 2030 mmr sdg projections (aka the mmr we hope to achieve for each country)
#Also, what does this return? Do we need to run this function multiple times
#In general, need help understanding this function
#is this global? since the next get sdg arr target is country-specific
```{r}
get_arr_sdg_target(mmr2015, live_birth_projections2030, 15)
```

#Calculate SDG ARR for each country Based on SDG MMR 
#From Pseudo Code
```{r}
get_arr_sdg_target_country(mmr2015, arr_sdg, 15)

#Other checks
#are the numbers I'm getting reasonable?

mmr2030_country <- get_mmr2030_sdg_projections(mmr_est_unrounded_pwider, arr_sdg)
##
arr_sdg_predictions <- get_arr_sdg_target_country(mmr2015, arr_sdg, 15)
mmr_est_unrounded_pwider %>% 
  cbind(arr_sdg_predictions) %>%
  rename(sdg_arr = `MMR2015`)

arr_sdg_predictions
#compare bau arr values
mmr_allcountries_projections(mmr_est_unrounded_pwider, 2010, 2017)

#Does this return correctly values? (B/c what about the adjustments to 140 that we made? Like, does this function know we're going from the original MMR to the adjusted one?)


get_mmr2030_sdg_projections(mmr2015, global_arr) %>% 
  cbind(get_mmr2030_sdg_projections(mmr2015, arr_sdg_predictions)) %>%
  rename(sdg_arr = `get_mmr2030_sdg_projctions(mmr2015, arr_sdg_predictions)`)

```

#(BAU) MMR Regional Summaries
- original parameters: mmr tibble, countries/regions tibble, live birth tibble
- UNHARDCODE specified years when they're the name of the columns
- UNHARDCODE the stuff inside the summarize function if possible, ie see if a column name can vary so the general formula can written just once (iterations?)
- returns: summary table of mmr by year for different regions and globally for a specified time range

#Consider making a version for this using SDG numbers
```{r}
#some data cleaning
countries_and_regions <- country_info %>% 
  left_join(regional_groupings, by = c("ISOCode" = "ISOCode")) %>%
  select(-c(`Country.x`)) 

#test
mmr_regional_global_summarize(mmr_est_unrounded_pwider, countries_and_regions, live_birth_projections2030)
```

#Recode of bau mmr summary function
```{r}
#function pt1 (global summaries only)
mrgs_pt1 <-
  function(mmr_pivotwider_tibble, countries_regions_tibble, birth_projections_tibble){
  mmr_df <- left_join(mmr_allcountries_projections(mmr_pivotwider_tibble, 2010, 2017) %>% select(-c(`name`, `arr`)), countries_regions_tibble, by = c("iso" = "ISOCode")) %>% left_join(birth_projections_tibble %>% select(`LocID`, `Births`), by = c("ISONum" = "LocID"))
  
nvars <- ncol(mmr_df) -  4
global_mmrs <- rep(NA, nvars) 
for (i in 2:nvars){ 
  global_mmrs[i] <- (sum(mmr_df[,i] * mmr_df$Births)) / (sum(mmr_df$Births)) 
} 

global <- na.omit(global_mmrs)
global
start_year = 2016
end_year = 2030
nyears <- end_year - start_year +1
years <- seq(start_year, end_year)
gm <- data.frame(years, global)
gm %>% 
  pivot_wider(names_from = `years`, values_from = `global`) %>%
      mutate(`sdg_1` = "Global") %>%
    select(c(16, seq(1, 15)))
  }

#example
mrgs_pt1(mmr_est_unrounded_pwider, countries_and_regions, live_birth_projections2030)

# #pt two of function in the works

md <- left_join(mmr_allcountries_projections(mmr_est_unrounded_pwider, 2010, 2017) %>% select(-c(`name`, `arr`)), countries_and_regions, by = c("iso" = "ISOCode")) %>% left_join(live_birth_projections2030 %>% select(`LocID`, `Births`), by = c("ISONum" = "LocID")) 

mrgs_regions <- function(mmr_pivotwider_tibble, countries_regions_tibble, birth_projections_tibble, sdg1_region) {
  #note that region must be in quotes
md_region <- filter(md, sdg_1 == sdg1_region)

nvars <- ncol(md_region) -  4
region_mmrs <- rep(NA, nvars) 
for (i in 2:nvars){ 
  region_mmrs[i] <- (sum(md_region[,i] * md_region$Births)) / (sum(md_region$Births)) 
} 

region_mmrs <- na.omit(region_mmrs)
region_mmrs
start_year = 2016
end_year = 2030
nyears <- end_year - start_year + 1
year <- seq(start_year, end_year)
region <- data.frame(year, region_mmrs)
region %>% 
  pivot_wider(names_from = `year`, values_from = `region_mmrs`) %>%
      mutate(`sdg_1` = sdg1_region) %>%
    select(c(16, seq(1, 15)))
}


##
region <- c() #what does this do? note the use of "region" above

region[1] <- "Australia and New Zealand"	
region[2] <- "Central Asia and Southern Asia"	
region[3] <- "Eastern Asia and South-eastern Asia"	
region[4] <- "Latin America and the Caribbean"	
region[5] <- "Northern America and Europe"	
region[6] <- "Oceania / Oceania excluding Australia and New Zealand"
region[7] <- "Sub-Saharan Africa"	
region[8] <- "Western Asia and Northern Africa"
region[9] <- "Global"
yr <- seq(2016, 2030)
st_tempcol2 <- rep(NA, 15)
summarytibble <- data.frame(yr, st_tempcol2) %>% 
  pivot_wider(names_from = `yr`, values_from = `st_tempcol2`) %>%
  mutate(sdg_1 = "Region")


Global = mrgs_pt1(mmr_est_unrounded_pwider, countries_and_regions, live_birth_projections2030) 
       ANZ = mrgs_regions(mmr_est_unrounded_pwider, countries_and_regions, live_birth_projections2030, region[1])
        CASA = mrgs_regions(mmr_est_unrounded_pwider, countries_and_regions, live_birth_projections2030, region[2])
       EASA = mrgs_regions(mmr_est_unrounded_pwider, countries_and_regions, live_birth_projections2030, region[3])
       LAC = mrgs_regions(mmr_est_unrounded_pwider, countries_and_regions, live_birth_projections2030, region[4])
       NAE = mrgs_regions(mmr_est_unrounded_pwider, countries_and_regions, live_birth_projections2030, region[5])
       OOeANZ = mrgs_regions(mmr_est_unrounded_pwider, countries_and_regions, live_birth_projections2030, region[6])
       SSA = mrgs_regions(mmr_est_unrounded_pwider, countries_and_regions, live_birth_projections2030, region[7])
       WANA = mrgs_regions(mmr_est_unrounded_pwider, countries_and_regions, live_birth_projections2030, region[8])

region_projections <- list(Global, ANZ, CASA, EASA, LAC, NAE, OOeANZ, 
                           SSA, WANA)


summary_table <- do.call(rbind, region_projections)
rename(summary_table, region = "sdg_1")
summary_table

#we need to pivotwider last b/c first we'll make the mmr summary results into columns to use a for loop

```