---
title: "WHO MMR Working Notebook"
output: html_notebook
---

#Read-in Data
```{r}
library(tidyr)
library(tidyverse)
library(readxl)
library(devtools)
library(usethis)

Rfiles <- list.files(file.path(paste0(getwd(),"/R/")), ".R")
Rfiles <- Rfiles[grepl(".R", Rfiles)]
sapply(paste0(paste0(getwd(),"/R/"), Rfiles), source)

country_info <- read_excel("country list_ 26 March 2019.xlsx")
mmr_est_unrounded <- read.csv("mmr_unrounded.csv")
live_birth_projections <- read_excel("wpp2019_Births-TFR-GFR-Female1549.xlsx")
regional_groupings <- read.csv("regional_groupings_20190910_la_ac.csv")

```

#Data Cleaning
```{r}
country_info <- country_info %>%
  select(`Code`, `ISO_Numeric_Code_CODE`, `Title`) %>%
  rename(ISOCode = `Code`, ISONum = `ISO_Numeric_Code_CODE`)
  
mmr_est_unrounded <- mmr_est_unrounded %>%
  filter(`bound` == "point", `year` >= 2010) %>%
  select(-c(`q`, `perc`, `bound`)) %>%
  mutate("MMR" = `value` * 100000)

mmr_est_unrounded_pwider <- mmr_est_unrounded %>%
  select(-c(`value`)) %>%
  pivot_wider(names_from = `year`, values_from = `MMR`)
  
regional_groupings <- regional_groupings %>%
  select(`ISOCode`, `Country.x`, `sdg_1`)

live_birth_projections2030 <- live_birth_projections %>% 
  filter(Year == 2030) %>%
  select(-c(`Year`)) %>%
  rename(name = 'Location') 
```

#Calc BAU ARR
- original parameters: mmr tibble
- UN-HARDCODE specified years when they're the name of the columns
- returns: tibble w/ added observed ARR for specified time period

```{r}
calc_bau_arr(mmr_est_unrounded_pwider, 2010, 2017)

#new function only has bau arr and iso code columns
cba(mmr_est_unrounded_pwider)


#Question:
# the function considers the length of the time period to be 7 years but it's actually 8 years b/c we're counting 2010 in the period
```


#MMR Projections (for one country)
- original parameters: mmr tibble, time1, time2, country_iso_code
- UNHARDCODE the number 2015 and the column name
- returns: a single column tibble w/ all the projected MMRs

```{r}
mmr_country_projections(mmr_est_unrounded_pwider, 2016, 2030, 2010, 2017, iso_code = "AFG")
    
#Recoded
mcp(mmr_est_unrounded_pwider, "AFG", 2016, 2030)

```

#MMR Projections (all countries)
- original parameters: mmr tibble, time1, time2
- UNHARDCODE column names
- RENAME "col" and "namecountry2" to something more intuitive
- returns: a single column tibble w/ all the projected MMRs
- is there a way to remove the column numbers of the final tibble?

```{r}
mmr_allcountries_projections(mmr_est_unrounded_pwider, 2010, 2017, 2016, 2030)

#(mmr_pivotwider_tibble, bau_start_year, bau_end_year, mmr_start_year, mmr_end_year)

#new function does not have arr column
macp(mmr_est_unrounded_pwider, 2016, 2030)
```

#SDG MMR Calculation, Categorization, and Adjustment
#This function currently calculates the MMR of 2030 based on a global arr from a simple calculation
- original parameters: mmr tibble, single year, global arr
- UNHARDCODE the number 2015 and the column name; the 2010 and 2017 in calc_bau_arr() function call
- SOLVE the issues w/ multiple tibbles and the resulting dependencies
- returns: tibble w/ countries'observed arr, mmr projections during time period, 2030 projection, each country's MMR category, each country's adjusted SDG goal
- Do we want this to simply add a column to the tibble passed thru asa an argument? Or just the calculated SDG MMR? Or maybe that plus the country codes/names? 

```{r}
global_arr <- mean(cba(mmr_est_unrounded_pwider)$`arr`)

mmr2015 <- mmr_est_unrounded_pwider %>% 
  rename(MMR2015 = `2015`) %>%
  select(`MMR2015`)

get_mmr_sdg_projections(mmr2015, global_arr, 15)

```

#Squared Diff
- for optimization purposes in conjunction w/ function above
```{r}
squared_diff(global_arr, mmr2015, live_birth_projections2030, 15)
```

#Get ARR SDG Target 
- original parameters: mmr2015_country (vector w/ MMR for 2015 for all countries), births2030_country (vector w/ births for 2030 for all countries), nproject (number of years to project (2030 - 2015))
- UNHARDCODE the specified years each argument holds (eg make it possible for users to use different years for projection bases and different year ranges, etc)
- consider making the values that the initial arguments hold different and actually calculated in the function body
- returns: tibble w/ global(?) arr sdg target for the cumulative period of 15 years
#Pseudo code
#This should feed into the 2030 mmr sdg projections (aka the mmr we hope to achieve for each country)
#Also, what does this return? Do we need to run this function multiple times
#In general, need help understanding this function
#is this global? since the next get sdg arr target is country-specific
```{r}
get_arr_sdg_target(mmr2015, live_birth_projections2030, 15)
arr_sdg <- 0.05603232
```

#Calculate SDG ARR for each country Based on SDG MMR 
#From Pseudo Code
```{r}
get_arr_sdg_target_country(mmr2015, arr_sdg, 15)

# #Other checks
# #are the numbers I'm getting reasonable?
# 
# mmr_yr_n_country <- get_mmr_sdg_projections(mmr_est_unrounded_pwider, arr_sdg, 15)
# ##
# arr_sdg_predictions <- get_arr_sdg_target_country(mmr2015, arr_sdg, 15)
# mmr_est_unrounded_pwider %>% 
#   cbind(arr_sdg_predictions) %>%
#   rename(sdg_arr = `MMR2015`)
# 
# arr_sdg_predictions
# #compare bau arr values
# mmr_allcountries_projections(mmr_est_unrounded_pwider, 2010, 2017)

```

#(BAU) MMR Regional Summaries
- original parameters: mmr tibble, countries/regions tibble, live birth tibble
- returns: summary table of mmr by year for different regions and globally for a specified time range

#Consider making a version for this using SDG numbers
```{r}
#some data cleaning
countries_and_regions <- country_info %>% 
  left_join(regional_groupings, by = c("ISOCode" = "ISOCode")) %>%
  select(-c(`Country.x`)) 

#test
mmr_regional_global_summarize(mmr_est_unrounded_pwider, countries_and_regions, live_birth_projections2030)

#Recoded version

#data cleaning
#mmr, countries and regions, live births combined tibble
md <- left_join(mmr_allcountries_projections(mmr_est_unrounded_pwider, 2010, 2017, 2016, 2030) %>% select(-c(`name`, `arr`)), countries_and_regions, by = c("iso" = "ISOCode")) %>% left_join(live_birth_projections2030 %>% select(`LocID`, `Births`), by = c("ISONum" = "LocID")) 

#example
bau_mmr_regional_projection_summaries(mmr_est_unrounded_pwider, countries_and_regions, live_birth_projections2030, 2010, 2017, 2016, 2030)
```


#Summary Graph
```{r}
bau_mmr_regional_global_graph(mmr_est_unrounded_pwider, countries_and_regions, live_birth_projections2030, 2010, 2017, 2016, 2030)

```



#Various tests
```{r}
#does the calculated sdg arr lead to a global mmr of 70?

  #Global MMR using SDG ARR (avg of all countries)
  (sum(get_mmr_sdg_projections(mmr2015, sum(arr_sdg_predictions) / 185, 15) * md$Births)) / (sum(md$Births)) #68.754

  #Global MMR using BAU ARR (avg of all countries)
  bau_mmr_proj <- mmr_est_unrounded_pwider$`2015` * exp(-(rep(0.0273, 185)) * (2030-2015))
   (sum(bau_mmr_proj * md$Births)) / (sum(md$Births)) #166.529
  #(is 170.117 by summary table calculations (using all countries' invidual arrs) above)

  
  #Global MMR using ARR = 2.9%
  bau_mmr_proj2 <- mmr_est_unrounded_pwider$`2015` * exp(-(rep(0.029, 185)) * (2030-2015))
  (sum(bau_mmr_proj2 * md$Births)) / (sum(md$Births)) #162.336

  
#
#(sum(get_mmr_sdg_projections(mmr2015, 0.04, 15) * md$Births)) / (sum(md$Births)) #80.325


```

