---
title: "Function Notebook"
output: html_notebook
---
#Contains all functions as they were written for final project last semester

```{r}
library(tidyr)
library(tidyverse)
library(readxl)
library(dplyr)
library(devtools)
library(usethis)
library(ggplot2)

country_info <- read_excel("country list_ 26 March 2019.xlsx")
mmr_est_unrounded <- read.csv("mmr_unrounded.csv")
live_birth_projections <- read_excel("wpp2019_Births-TFR-GFR-Female1549.xlsx")
regional_groupings <- read.csv("regional_groupings_20190910_la_ac.csv")

```

#Data Cleaning
```{r}
country_info <- country_info %>%
  select(`Code`, `ISO_Numeric_Code_CODE`, `Title`) 
  
mmr_est_unrounded <- mmr_est_unrounded %>%
  filter(`bound` == "point", `year` >= 2010) %>%
  select(-c(`q`, `perc`, `bound`)) %>%
  mutate("MMR" = `value`*100000)

mmr_est_unrounded_pwider <- mmr_est_unrounded %>%
  select(-c(`value`)) %>%
  pivot_wider(names_from = `year`, values_from = `MMR`)
  
regional_groupings <- regional_groupings %>%
  select(`ISOCode`, `Country.x`, `sdg_1`)

live_birth_projections2 <- live_birth_projections %>% 
  filter(Year == 2030)
colnames(live_birth_projections2)[colnames(live_birth_projections2)=="Location"] <- "name"
```

#BAU ARR Function
Calculates the business as usual ARR for all countries (adds column to provided tibble)
```{r}
bau_arr <- function(mmr_est_unrounded_pwider, t1, t2) {
  length = t2-t1
  arr = -1/(length)*log(mmr_est_unrounded_pwider$`2017`/mmr_est_unrounded_pwider$`2010`)
  arr_percent=arr*100
  return(cbind(mmr_est_unrounded_pwider,arr,arr_percent))
}

bau_arr(mmr_est_unrounded_pwider, 2010, 2017)
```

#MMR Projection Function (One Country)
Calculates the projected MMRs from 2016 to 2030 for a single, specified country
```{r}
mmr_country_projections<-function(mmr_est_unrounded_pwider, t1 = 2010, t2 = 2017, iso_code){

  years <- seq(2016, 2030)
  nyears <- length(years)
  dat2<- bau_arr(mmr_est_unrounded_pwider, t1, t2) %>% filter(iso == iso_code)
  projection <- rep(NA,nyears)
  projection <- dat2$`2015` *exp(-(dat2$`arr`) * (years-2015))
  return(projection)
}

mmr_country_projections(mmr_est_unrounded_pwider, 2010, 2017, iso_code = "AFG")
```

#MMR Projection Function (All Countries)
Calculates the projected MMRs from 2016 to 2030 for all countries
```{r}
mmr_allcountries_projections<-function(mmr_est_unrounded_pwider, t1, t2){
  df3  <- NULL
  for(i in 1:nrow(bau_arr(mmr_est_unrounded_pwider, t1, t2))){
    col <- mmr_country_projections(mmr_est_unrounded_pwider, iso_code = bau_arr(mmr_est_unrounded_pwider, t1, t2)$iso[i])
    df3 <- rbind(df3, col)
    colnames(df3) <- c("2016", "2017", "2018", "2019", "2020", "2021", "2022", "2023", "2024", "2025", "2026", "2027", "2028", "2029", "2030")
  }

  namecountry2<- select(bau_arr(mmr_est_unrounded_pwider, t1, t2), "iso" , "name", "arr")
  mmrprojcountry2<- cbind(namecountry2, df3)
  return(mmrprojcountry2)
}

mmr_allcountries_projections(mmr_est_unrounded_pwider, 2010, 2017)
```

#SDG MMR Calculation, Categorization, and Adjustment
Calculates the projected MMR for a specified year based on a global_arr, then:
- categorizes the country's MMR as high or acceptable 
- readjusts the high MMR countries' MMR to meet the SDG requirements
```{r}
global_arr<- mean(bau_arr(mmr_est_unrounded_pwider, 2010, 2017)$`arr`)

SDG_MMR_2030 <- function(mmr_est_unrounded_pwider, t, global_arr) {
  projection <- bau_arr(mmr_est_unrounded_pwider, 2010, 2017)$`2015` * exp(-global_arr * (t- 2015))
  MMRtest <- cbind(mmr_allcountries_projections(mmr_est_unrounded_pwider, 2010, 2017), projection)
  MMRtest1 <- MMRtest %>% mutate(category=ifelse(projection>140,"high mmr","acceptable mmr")) %>% mutate(newprojection=ifelse(projection>140,140,projection))
  return(MMRtest1)
}

SDG_MMR_2030(mmr_est_unrounded_pwider, 2030, global_arr)
```

#Regional MMR Summary Table
Calculates the projected MMRs from 2016 to 2030 based on region
```{r}
#some data cleaning
colnames(country_info)[colnames(country_info)== "Code"] <- "ISOCode"

countries_and_regions <- country_info %>% 
  left_join(regional_groupings, by = c("ISOCode" = "ISOCode")) %>%
  select(-c(`Country.x`))

colnames(countries_and_regions)[colnames(countries_and_regions)== "ISO_Numeric_Code_CODE"] <- "ISONum"

#function
mmr_regional_global_summarize <-
  function(mmr_est_unrounded_pwider, countries_and_regions, live_birth_projections2){
    mmr_df <- left_join(mmr_allcountries_projections(mmr_est_unrounded_pwider, 2010, 2017), countries_and_regions, by = c("iso" = "ISOCode")) %>% left_join(live_birth_projections2, by = c("ISONum" = "LocID"))
    global <- mmr_df %>%
      summarise(`MMR in 2016` = (sum(`2016`*`Births`))/(sum(`Births`)),
                `MMR in 2017` = (sum(`2017`*`Births`))/(sum(`Births`)),
                `MMR in 2018` = (sum(`2018`*`Births`))/(sum(`Births`)),
                `MMR in 2019` = (sum(`2019`*`Births`))/(sum(`Births`)),
                `MMR in 2020` = (sum(`2020`*`Births`))/(sum(`Births`)),
                `MMR in 2021` = (sum(`2021`*`Births`))/(sum(`Births`)),
                `MMR in 2022` = (sum(`2022`*`Births`))/(sum(`Births`)),
                `MMR in 2023` = (sum(`2023`*`Births`))/(sum(`Births`)),
                `MMR in 2024` = (sum(`2024`*`Births`))/(sum(`Births`)),
                `MMR in 2025` = (sum(`2025`*`Births`))/(sum(`Births`)),
                `MMR in 2026` = (sum(`2026`*`Births`))/(sum(`Births`)),
                `MMR in 2027` = (sum(`2027`*`Births`))/(sum(`Births`)),
                `MMR in 2028` = (sum(`2028`*`Births`))/(sum(`Births`)),
                `MMR in 2029` = (sum(`2029`*`Births`))/(sum(`Births`)),
                `MMR in 2030` = (sum(`2030`*`Births`))/(sum(`Births`))) %>%
      mutate(`sdg_1` = "Global") %>%
      select(`sdg_1`, `MMR in 2016`, `MMR in 2017`, `MMR in 2018`, `MMR in 2019`, `MMR in 2020`, `MMR in 2021`, `MMR in 2022`, `MMR in 2023`, `MMR in 2024`, `MMR in 2025`, `MMR in 2026`, `MMR in 2027`, `MMR in 2028`, `MMR in 2029`, `MMR in 2030`)
    regional <- mmr_df %>%
      group_by(sdg_1) %>%
      summarise(`MMR in 2016` = (sum(`2016`*`Births`))/(sum(`Births`)),
                `MMR in 2017` = (sum(`2017`*`Births`))/(sum(`Births`)),
                `MMR in 2018` = (sum(`2018`*`Births`))/(sum(`Births`)),
                `MMR in 2019` = (sum(`2019`*`Births`))/(sum(`Births`)),
                `MMR in 2020` = (sum(`2020`*`Births`))/(sum(`Births`)),
                `MMR in 2021` = (sum(`2021`*`Births`))/(sum(`Births`)),
                `MMR in 2022` = (sum(`2022`*`Births`))/(sum(`Births`)),
                `MMR in 2023` = (sum(`2023`*`Births`))/(sum(`Births`)),
                `MMR in 2024` = (sum(`2024`*`Births`))/(sum(`Births`)),
                `MMR in 2025` = (sum(`2025`*`Births`))/(sum(`Births`)),
                `MMR in 2026` = (sum(`2026`*`Births`))/(sum(`Births`)),
                `MMR in 2027` = (sum(`2027`*`Births`))/(sum(`Births`)),
                `MMR in 2028` = (sum(`2028`*`Births`))/(sum(`Births`)),
                `MMR in 2029` = (sum(`2029`*`Births`))/(sum(`Births`)),
                `MMR in 2030` = (sum(`2030`*`Births`))/(sum(`Births`)))
    rbind(global, regional) %>%
      rename(Region=sdg_1)
  }

mmr_regional_global_summarize(mmr_est_unrounded_pwider, countries_and_regions, live_birth_projections2)
```

#Plot Regional Summaries
```{r}
mmr_regional_global_summaries <-
mmr_regional_global_summarize(mmr_est_unrounded_pwider, countries_and_regions, live_birth_projections2)

mmr_graph <- mmr_regional_global_summaries %>%
  rename(`2016`="MMR in 2016") %>%
  rename(`2017`="MMR in 2017") %>%
  rename(`2018`="MMR in 2018") %>%
  rename(`2019`="MMR in 2019") %>%
  rename(`2020`="MMR in 2020") %>%
  rename(`2021`="MMR in 2021") %>%
  rename(`2022`="MMR in 2022") %>%
  rename(`2023`="MMR in 2023") %>%
  rename(`2024`="MMR in 2024") %>%
  rename(`2025`="MMR in 2025") %>%
  rename(`2026`="MMR in 2026") %>%
  rename(`2027`="MMR in 2027") %>%
  rename(`2028`="MMR in 2028") %>%
  rename(`2029`="MMR in 2029") %>%
  rename(`2030`="MMR in 2030") %>%
pivot_longer(2:16,names_to="period",values_to="mmr")
mmr_graph %>%
ggplot() + geom_point(aes(x=period,y=mmr, color=Region)) +
  geom_line(mapping = aes(x=period, y=mmr, group=Region, color = Region))
```
