---
title: "WHO MMR MD 4_24_20"
author: "Li Shandross"
date: "4/24/2020"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This document serves as documentation for the package WHOmmr that aims to help analyze the data collected by the World Health Organization (WHO) on Maternal Mortality Ratio (MMR) around the world. WHO defines MMR as "the number of maternal deaths per 100,000 live births." WHOmmr is a collection of functions that work on estimates of maternal mortality between 2000 and 2017 (though we are focusing on 2010 to 2017) to gain insight about the projected MMR for every country between 2016 and 2030. In particular, this package is meant to assist WHO in the achieving the standard development goal (SDG) of a global MMR of 70 by 2030, with no single country having an MMR above 140, to reduce inequality in this sector. 

We compare MMR projections following the observed, Business As Usual (BAU) trends with these SDG projections that WHO hopes to achieve. The Average Rate of Reduction (ARR) is the average annual reduction in MMR a particular country or region experiences, and this ARR is used to calculate the desired projections.  

Each function below has more information about its purpose. 

#Read-in Data
```{r}
library(tidyr)
library(tidyverse)
library(readxl)
library(dplyr)
library(devtools)
library(usethis)
library(knitr)

load_all("WHOmmr")

country_info <- read_excel("country list_ 26 March 2019.xlsx")
mmr_est_unrounded <- read.csv("mmr_unrounded.csv")
live_birth_projections <- read_excel("wpp2019_Births-TFR-GFR-Female1549.xlsx")
regional_groupings <- read.csv("regional_groupings_20190910_la_ac.csv")

```  



#Data Cleaning
```{r}
country_info <- country_info %>%
  select(`Code`, `ISO_Numeric_Code_CODE`, `Title`) %>%
  rename(ISOCode = `Code`, ISONum = `ISO_Numeric_Code_CODE`)
  
mmr_est_unrounded <- mmr_est_unrounded %>%
  filter(`bound` == "point", `year` >= 2010) %>%
  select(-c(`q`, `perc`, `bound`)) %>%
  mutate("MMR" = `value` * 100000)

mmr_est_unrounded_pwider <- mmr_est_unrounded %>%
  select(-c(`value`)) %>%
  pivot_wider(names_from = `year`, values_from = `MMR`)
  
regional_groupings <- regional_groupings %>%
  select(`ISOCode`, `Country.x`, `sdg_1`)

live_birth_projections2030 <- live_birth_projections %>% 
  filter(Year == 2030) %>%
  select(-c(`Year`)) %>%
  rename(name = 'Location') 
```

#Calc BAU ARR
Calculates the observed (BAU) ARR of each country based on the observed ARR from time1 to time2. Base Equation: ARR (for period t1 to t2) = -1/(t2-t1)*log(MMR(t2)/MMR(t1))

Tests:
1. Check inputs (stop if inputs for arguments are not as expected), both in terms of data type and in terms of if the numerical arguments are within an acceptable range. 
2. Test to see if correct values are returned (using toy data) based on base equation in function.
```{r}
calc_bau_arr_tibble <- calc_bau_arr(mmr_est_unrounded_pwider, 3, 10)
kable(calc_bau_arr_tibble[1:5, ])
```


#MMR Projections (for one country)
Calculates the MMR projections for one country for a specified period using data from baseyear 2015 and the country's respective BAU ARR. The single country is specified by using its assigned ISO code.  
Base Equation: MMR(t) = MMR(2015)*exp(-ARR * (t-2015))

Tests:
1. Check inputs (stop if inputs for arguments are not as expected), both in terms of data type and in terms of if the numerical arguments are within an acceptable range.
2. Test to see if correct values are returned (using toy data) based on base equation in function.
3. Check that this function's call to calc_bau_arr produces the correct results. 
```{r}
bau_mmr_single_country_proj(mmr_est_unrounded_pwider, "AFG", 3, 10, 2016, 2030)
  
```


#MMR Projections (all countries)
Calculates the MMR projections for all countries for a specified period using data from baseyear 2015 and each country's respective BAU ARR. 
Base Equation: MMR(t) = MMR(2015)*exp(-ARR * (t-2015))

Tests:
1. Check inputs (stop if inputs for arguments are not as expected), both in terms of data type and in terms of if the numerical arguments are within an acceptable range.
2. Test to see if correct values are returned (using toy data) based on base equation in function.
3. Check that this function's call to bau_mmr_single_country_proj (and thus calc_bau_arr, as well) produces the correct results. 
```{r}
bau_mmr_proj_tibble <- bau_mmr_all_countries_proj(mmr_est_unrounded_pwider,3, 10, 2016, 2030)
kable(bau_mmr_proj_tibble[1:6, ])
```



#SDG MMR Calculation, Categorization, and Adjustment
Calculates the MMR projections for all using a fixed value of the ARR for a single year t > 2015 using baseyear 2015, then adjusts the projections to be less than or equal to 140, as specified by the WHO's SDG.
Base Equation: MMR(t) = MMR(2015)*exp(-ARR * (t-2015))
For countries with mmr_target2030 > 140, replace mmr_target by 140, mmr_target_final = ifelse(mmr_target2030 > 140, 140, mmr_target)

Tests:
1. Check inputs (stop if inputs for arguments are not as expected), both in terms of data type and in terms of if the numerical arguments are within an acceptable range.
2. Test to see if correct values are returned (using toy data) based on base equation in function.
```{r}
global_arr <- mean(calc_bau_arr(mmr_est_unrounded_pwider, 3, 10)$`arr`)

mmr2015 <- mmr_est_unrounded_pwider %>% 
  rename(MMR2015 = `2015`) %>%
  select(`MMR2015`)

mmr_sdg_proj <- get_mmr_sdg_proj(mmr2015, global_arr, 15)
kable(mmr_sdg_proj[1:5, ])
```


#Squared Diff
Calculates the squared difference of the global mmr of a specified year and the SDG goal of a global MMR of 70.
Base Equation: (global_mmr - 70)^2

Tests:
1. Check inputs (stop if inputs for arguments are not as expected), both in terms of data type and in terms of if the numerical arguments are within an acceptable range.
2. Test to see if correct values are returned (using toy data) based on base equation in function.
3. Check that this function's call to get_mmr_sdg_proj produces the correct results. 
```{r, warning=F}
squared_diff(global_arr, mmr2015, live_birth_projections2030, 15)
```


#Get ARR SDG Target 
Calculates the target ARR needed to achieve WHO's SDG goal of a global MMR of 70, with no country with an MMR above 140, by 2030.

Tests:
1. Check inputs (stop if inputs for arguments are not as expected), both in terms of data type and in terms of if the numerical arguments are within an acceptable range.
2. Test to see if correct values are returned (using toy data) based on base equation in the function.
3. Check that this function's call to squared_diff (and thus get_mmr_sdg_proj, as well) produces the correct results.
```{r, warning=F}
get_arr_sdg_target(mmr2015, live_birth_projections2030, 15)
sdg_arr <- get_arr_sdg_target(mmr2015, live_birth_projections2030, 15)$minimum
```

#Calculate SDG ARR for each country Based on SDG MMR 
Calculates the specific ARR for each country needed to achieve WHO's SDG goal of a global MMR of 70, with no country with an MMR above 140, by 2030. Each country-specific ARR is based on the single target SDG ARR calculated. 

Base Equation: -1/nproject * log(mmr_sdg_projections_using_sdg_arr/2015_mmr_all_countries)

Tests:
1. Check inputs (stop if inputs for arguments are not as expected), both in terms of data type and in terms of if the numerical arguments are within an acceptable range.
2. Test to see if correct values are returned (using toy data) based on base equation in the function.
3. Check that this function's call to get_mmr_sdg_proj and get_arr_sdg_proj (and thus squared_diff, as well) produces the correct results.
```{r, warning=F}
 arr_sdg_target_country <- calc_sdg_arr(mmr_est_unrounded_pwider, mmr2015, live_birth_projections2030, 15)
kable(arr_sdg_target_country[1:5, ])
```


#(BAU) MMR Regional Summaries
1) Produces a table of all the projected BAU MMRs by SDG region for a specified period.
2) Produces a line graph of all the projected BAU MMRs by SDG region.

Base Equation: MMR(region) = sum(region_mmr_proj*region_total_births) / sum(region_total_births)

Tests:
1. Check inputs (stop if inputs for arguments are not as expected), both in terms of data type and in terms of if the numerical arguments are within an acceptable range.
2. Test to see if correct values are returned (using toy data) based on base equation in the function.
3. Check that this function's call to get_mmr_sdg_proj and get_arr_sdg_proj (and thus squared_diff, as well) produces the correct results.
```{r, warning=F}
#some data cleaning
countries_and_regions <- country_info %>% 
  left_join(regional_groupings, by = c("ISOCode" = "ISOCode")) %>%
  select(-c(`Country.x`)) 

#Part 1: Table
regional_proj_summaries <- bau_mmr_regional_projection_summaries(mmr_est_unrounded_pwider, countries_and_regions, live_birth_projections2030, 3, 10, 2016, 2030)
knitr::kable(regional_proj_summaries)

#Part 2: Graph
bau_mmr_regional_global_graph(mmr_est_unrounded_pwider, countries_and_regions, live_birth_projections2030, 3, 10, 2016, 2030)
 
```

#SDG MMR Projections
Calculates the SDG MMR projections for all countries for a specified period using data from baseyear 2015 and each country's respective BAU ARR. 

Base Equation: MMR(t) = MMR(2015)*exp(-ARR * (t-2015))

Tests:
1. Check inputs (stop if inputs for arguments are not as expected), both in terms of data type and in terms of if the numerical arguments are within an acceptable range.
2. Test to see if correct values are returned (using toy data) based on base equation in the function.
3. Check that this function's call to sdg_mmr_single_country_proj produces the correct results.
```{r, warning=F}
arr_tibble <- left_join(calc_bau_arr(mmr_est_unrounded_pwider, 3, 10), calc_sdg_arr(mmr_est_unrounded_pwider, mmr2015, live_birth_projections2030, 15), by = c("iso" = "iso")) 
  arr_tibble <- rename(arr_tibble, `sdg arr` = sdg_arr, `bau arr` = arr)

sdg_mmr_single_country_proj(mmr_est_unrounded_pwider, arr_tibble, "AFG", 2016, 2030)

sdg_mmr_proj <- sdg_mmr_allcountries_proj(mmr_est_unrounded_pwider, arr_tibble, 2016, 2030) %>%
  select(-c("bau arr"))
 
kable(sdg_mmr_proj[1:6, ])
kable(bau_mmr_proj_tibble[1:6, ]) #maybe delete?
  
```

