---
title: "WHO MMR MD 4_18_20"
author: "Li Shandross"
date: "4/18/2020"
output: github_document
---

This document serves as documentation for the package [PACKAGE NAME] that aims to help analyze the data collected by the World Health Organization (WHO) on Maternal Mortality Ratio (MMR) around the world. WHO defines MMR as "the number of maternal deaths per 100,000 live births." [PACKAGE NAME] is a collection of functions that work on estimates of maternal mortality between 2000 and 2017 (though we are focusing on 2010 to 2017) to gain insight about the projected MMR for every country between 2016 and 2030. In particular, this package is meant to assist WHO in the achieving the standard development goal (SDG) of a global MMR of 70 by 2030, with no single country having an MMR above 140, to reduce inequality in this sector. 

We compare MMR projections following the observed, Business As Usual (BAU) trends with these SDG projections that WHO hopes to achieve. The Average Rate of Reduction (ARR) is the average annual reduction in MMR a particular country or region experiences, and this ARR is used to calculate the desired projections.  

Each function below has more information about its purpose. 


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Read-in Data
```{r}
library(tidyr)
library(tidyverse)
library(readxl)
library(devtools)
library(usethis)
library(knitr)

Rfiles <- list.files(file.path(paste0(getwd(),"/R/")), ".R")
Rfiles <- Rfiles[grepl(".R", Rfiles)]
sapply(paste0(paste0(getwd(),"/R/"), Rfiles), source)

country_info <- read_excel("country list_ 26 March 2019.xlsx")
mmr_est_unrounded <- read.csv("mmr_unrounded.csv")
live_birth_projections <- read_excel("wpp2019_Births-TFR-GFR-Female1549.xlsx")
regional_groupings <- read.csv("regional_groupings_20190910_la_ac.csv")
```


#Data Cleaning
```{r}
country_info <- country_info %>%
  select(`Code`, `ISO_Numeric_Code_CODE`, `Title`) %>%
  rename(ISOCode = `Code`, ISONum = `ISO_Numeric_Code_CODE`)
  
mmr_est_unrounded <- mmr_est_unrounded %>%
  filter(`bound` == "point", `year` >= 2010) %>%
  select(-c(`q`, `perc`, `bound`)) %>%
  mutate("MMR" = `value` * 100000)

mmr_est_unrounded_pwider <- mmr_est_unrounded %>%
  select(-c(`value`)) %>%
  pivot_wider(names_from = `year`, values_from = `MMR`)
  
regional_groupings <- regional_groupings %>%
  select(`ISOCode`, `Country.x`, `sdg_1`)

live_birth_projections2030 <- live_birth_projections %>% 
  filter(Year == 2030) %>%
  select(-c(`Year`)) %>%
  rename(name = 'Location') 
```

#Calc BAU ARR
Calculates the observed (BAU) ARR of each country based on the observed ARR from time1 to time2. Base Equation: ARR (for period t1 to t2) = -1/(t2-t1)*log(MMR(t2)/MMR(t1))

```{r}
calc_bau_arr_tibble <- calc_bau_arr(mmr_est_unrounded_pwider, 2010, 2017)
kable(calc_bau_arr_tibble[1:5, ])

#new function only has bau arr and iso code columns
cba_tibble <- cba(mmr_est_unrounded_pwider)
kable(cba_tibble[1:5, ])

```


#MMR Projections (for one country)
Calculates the MMR projections for one country for a specified period using data from baseyear 2015 and the country's respective BAU ARR. The single country is specified by using its assigned ISO code.  
Base Equation: MMR(t) = MMR(2015)*exp(-ARR * (t-2015))
```{r}
mmr_country_projections(mmr_est_unrounded_pwider, 2016, 2030, 2010, 2017, iso_code = "AFG")
    
#Recoded
mcp(mmr_est_unrounded_pwider, "AFG", 2016, 2030)
```


#MMR Projections (all countries)
Calculates the MMR projections for all countries for a specified period using data from baseyear 2015 and each country's respective BAU ARR. 
Base Equation: MMR(t) = MMR(2015)*exp(-ARR * (t-2015))
```{r}
mmr_allcountries_proj_tibble <- mmr_allcountries_projections(mmr_est_unrounded_pwider, 2010, 2017, 2016, 2030)
kable(mmr_allcountries_proj_tibble[1:6, ])

#new function does not have arr column
macp_tibble <- macp(mmr_est_unrounded_pwider, 2016, 2030)
kable(macp_tibble[1:6, ])
```


#SDG MMR Calculation, Categorization, and Adjustment
Calculates the MMR projections for all using a fixed value of the ARR for a single year t > 2015 using baseyear 2015, then adjusts the projections to be less than or equal to 140, as specified by the WHO's SDG.
Base Equation: MMR(t) = MMR(2015)*exp(-ARR * (t-2015))
For countries with mmr_target2030 > 140, replace mmr_target by 140, mmr_target_final = ifelse(mmr_target2030 > 140, 140, mmr_target)
```{r}
global_arr <- mean(cba(mmr_est_unrounded_pwider)$`arr`)

mmr2015 <- mmr_est_unrounded_pwider %>% 
  rename(MMR2015 = `2015`) %>%
  select(`MMR2015`)

mmr_sdg_proj <- data.frame(get_mmr_sdg_projections(mmr2015, global_arr, 15))
#make data.frame in function instead?
kable(mmr_sdg_proj[1:5, ])
```


#Squared Diff
Calculates the squared difference of...
```{r}
squared_diff(global_arr, mmr2015, live_birth_projections2030, 15)
```


#Get ARR SDG Target 
```{r}
get_arr_sdg_target(mmr2015, live_birth_projections2030, 15)
arr_sdg <- 0.05603232
```


#Calculate SDG ARR for each country Based on SDG MMR 
#From Pseudo Code
```{r}
arr_sdg_target_country <- get_arr_sdg_target_country(mmr2015, arr_sdg, 15)
kable(arr_sdg_target_country[1:5, ])
```


#(BAU) MMR Regional Summaries
1) Produces a table of all the projected BAU MMRs by SDG region for a specified period.
2) Produces a line graph of all the projected BAU MMRs by SDG region.
```{r}
#some data cleaning
countries_and_regions <- country_info %>% 
  left_join(regional_groupings, by = c("ISOCode" = "ISOCode")) %>%
  select(-c(`Country.x`)) 

#example
regional_proj_summaries <- bau_mmr_regional_projection_summaries(mmr_est_unrounded_pwider, countries_and_regions, live_birth_projections2030, 2016, 2030)
knitr::kable(regional_proj_summaries)

#Graphical Representation
bau_mmr_regional_global_graph(mmr_est_unrounded_pwider, countries_and_regions, live_birth_projections2030, 2016, 2030)

```


#Various tests
```{r}
md <- left_join(macp(mmr_est_unrounded_pwider, 2016, 2030) %>% select(-c(`name`)), countries_and_regions, by = c("iso" = "ISOCode")) %>% left_join(live_birth_projections2030 %>% select(`LocID`, `Births`), by = c("ISONum" = "LocID"))
    

arr_sdg_predictions <- get_arr_sdg_target_country(mmr2015, arr_sdg, 15)

  #Global MMR using SDG ARR (avg of all countries)
  (sum(get_mmr_sdg_projections(mmr2015, sum(arr_sdg_predictions) / 185, 15) * md$Births)) / (sum(md$Births)) #68.754

  #Global MMR using BAU ARR (avg of all countries)
  bau_mmr_proj <- mmr_est_unrounded_pwider$`2015` * exp(-(rep(0.0273, 185)) * (2030-2015))
   (sum(bau_mmr_proj * md$Births)) / (sum(md$Births)) #166.529
  #(is 170.117 by summary table calculations (using all countries' invidual arrs) above)

  
  #Global MMR using ARR = 2.9%
  bau_mmr_proj2 <- mmr_est_unrounded_pwider$`2015` * exp(-(rep(0.029, 185)) * (2030-2015))
  (sum(bau_mmr_proj2 * md$Births)) / (sum(md$Births)) #162.336
  
##
  
  ###

arr_tibble <- cbind(cba(mmr_est_unrounded_pwider), get_arr_sdg_target_country(mmr2015, arr_sdg, 15)) %>%
  rename(`sdg arr` = MMR2015, `bau arr` = arr)

#sdg_mmr_single_country_proj(mmr_est_unrounded_pwider, arr_tibble, "AFG", 2016, 2030)

sdg_mmr_proj <- sdg_mmr_allcountries_proj(mmr_est_unrounded_pwider, arr_tibble, 2016, 2030) %>%
  select(-c("bau arr"))
kable(sdg_mmr_proj[1:6, ])
        
kable(mmr_allcountries_proj_tibble[1:6, ])
  

```